### 2011-02-27 paper/paper.tex

\documentclass{article}
\newcommand{\osn}{\oldstylenums}
\title{TITLE HERE}
\author{Alec Story and Thomas Levine\\\small{avs\osn{38} and tkl\osn{22}}}

\begin{document}
\maketitle
\section*{Sentence Generation}

	\subsection*{Unsmoothed}

		\subsubsection*{Unigrams}

			\begin{quote}

			Bring in Take made the pass a Caesar in gold well leave \#

			Exeunt . hope , and Milan your A Will thy have Come the . .
			advantage , together your either to '\#

			with . ratified who of are in my year to natural I have the do to
			Master Turkish there wherein \#

			\end{quote}

		\subsubsection*{Bigrams}

			\begin{quote}

			Enter LYSANDER , and you lie dull actor on my speech , an ordinary
			pitch , Loyal and full of our tongues , see , That which physic to
			my horse for her youth and sealed bags Of base earth ; Though twenty
			several mistress will you have done To have found . \#

			Silence , who . \#

			The king , go away ! \#

			An heir ; for the likeness of eternal in a raven black , the sons ,
			whisper me have all those that thou wert not me welcome . \#

			\end{quote}

		\subsubsection*{Trigrams}

			\begin{quote}

			Re-enter BIONDELLO , LUCENTIO , and goodness never fearful . \#

			Gracious , my lad . \#

			How - traitor ? \#

			The Greeks upon advice , Hath alter'd that good wisdom Whereof I know
			by their christen names , Unless thou tell her so . \#

			\end{quote}

	\subsection*{Smoothed}

\end{document}