### 2011-02-28 paper/paper.tex

\documentclass{article}
\newcommand{\osn}{\oldstylenums}
\title{TITLE HERE}
\author{Alec Story and Thomas Levine\\\small{avs\osn{38} and tkl\osn{22}}}

\begin{document}
\maketitle
\section*{Sentence Generation}

	For sentence generation, we did not include our unknown processing, and do
	not convert the first occurrence of each word to an unknown token.

	Each subsection contains two examples from the Shakespeare corpus and two
	from War and Peace.  These corpora were chosen because they were determined
	at rough glance to be the easiest to parse and had the fewest problematic

	\subsection*{Unsmoothed}

		\subsubsection*{Unigrams}

			\begin{quote}

			Bring in Take made the pass a Caesar in gold well leave

			Exeunt . hope , and Milan your A Will thy have Come the . .
			advantage , together your either to '

			He fell other along " therefore , and had about He chief shorten now
			him life . something

			I had transference shrugged man with you The lit . to lift the her
			had he . and have the hand individuals dragging the , place been to
			Natasha in inside . revelry hide nation room beaming eager all of
			the the , remounts to

			\end{quote}

		\subsubsection*{Bigrams}

			\begin{quote}

			Enter LYSANDER , and you lie dull actor on my speech , an ordinary
			pitch , Loyal and full of our tongues , see , That which physic to
			my horse for her youth and sealed bags Of base earth ; Though twenty
			several mistress will you have done To have found .

			Silence , who .

			The horses ' familiar to the army , envy and looked questioningly :
			"You won't be thus depriving him of the estate with his will send me
			forty thousand men in choosing her governess kept saying good-by .

			I have begun to Tushin and yet further battles , the man in the
			defense of that temple , they were standing beside whom you seen in
			the battle of dry branches of meeting it has brought the smile Berg
			with a mocking , and even impossible to be long paced angrily .

			\end{quote}

		\subsubsection*{Trigrams}

			\begin{quote}

			How - traitor ?

			The Greeks upon advice , Hath alter'd that good wisdom Whereof I know
			by their christen names , Unless thou tell her so .

			No , but it can't be helped It happens to the first to arrive ,
			Princess , I think .

			Rostopchin , coming to Petersburg , look out at night his feet .

			\end{quote}

	\subsection*{Smoothed}

		\subsubsection*{Unigram}

			\begin{quote}

			th ! come Lear things of the I

			me . I

			they its intimate just such But There the them said shout
			irresistible always heard none with and of dinner had To

			day good-natured Secondly same , joyful on talking the appeared
			advantage . unreasoning men by

			\end{quote}

		\subsubsection*{Bigram}

			\begin{quote}

			Hector .

			I arrest thee here , which time , that his fault , and blows .

			But my wife should have nobody and ask the enemy was too .

			Say it's farther .

			\end{quote}

		\subsubsection*{Trigram}

			\begin{quote}

			She would have made fair work !

			Come , my lord , I'll bring you to please his Majesty , Herod of
			Jewry dare not say he is much abus'd with tears thou keep'st command
			.

			All that he had no news from the chase .

			If I decline the honor of the bushes .

			\end{quote}

	\section*{Arbitrary Precision}

		One of the two extensions we chose was to implement the probability and
		perplexity calculations using arbitrary precision arithmetic (we also
		implemented it using logarithmic arithmetic).  This can lead to slower
		processing, but in our profiling, most of the time was spent in
		dictionary manipulation rather than mathematics, and guarantees that no
		information is lost.

		We used python's +gmpy+ package to represent the probabilities as
		arbitrary rational numbers (python's integers are already
		arbitrary-precision), and +mpmath+ to perform the final exponentiation
		to calculate the perplexity (the $n$-th root).

		Our hypothesis was that the arbitrary precision math would reveal a
		small amount of error in computing the perplexity logarithmically.

	\section*{Arbitrary $n$-grams}

		Because the computations were general enough, we were able to easily
		abstract the code to produce count, smoothed count, and probability
		tables for arbitrary $n$-grams.  We also compute all of the smaller
		$n$-grams, which are required for generating or analyzing the beginning
		of a sequence of words, and for other computations.
\end{document}