### 2011-03-01 paper/paper.Rnw

\documentclass{article}
\newcommand{\osn}{\oldstylenums}
\title{Arbitrary Precision and Arbitrary $n$-Grams}
\author{Alec Story and Thomas Levine\\\small{avs\osn{38} and tkl\osn{22}}}

\newcommand\ngram{$n$-gram}
\newcommand\loga{logarithmic arithmetic}

<<echo=F>>=
source('analysis.R')
@

\begin{document}

\maketitle
\section*{Implementation}

	Our design is built around dictionaries for each $n$-gram we consider.  We
	break the input text into words and add sentence separators (although we do
	not break the sentences apart, so trigrams or higher can cross sentence
	boundaries), and then process the list of words, tallying the number of each
	$n$-gram seen in the dictionary.

	We also can optionally account for unknown words by replacing the first
	instance of each word with an unknown symbol; this is simply an additional
	processing step applied to the raw list of words.

	We compute all $n$-gram count dictionaries up to the desired $n$-gram
	length, and use these to compute the probabilities of each $n$-gram.  This
	dictionary of probabilities is used to construct sentences and to measure
	perplexity.  We either use logarithmic or arbitrary-precision arithmetic
	depending on what the user specifies.

	We implemented Good-Turing smoothing, which adjusts the $n$-gram counts
	before they are turned into probabilities.  Like unknown-handling, this is
	simply a pipeline processing step.

	Sentence generation begins with a beginning of sentence symbol, and then
	chooses words from the $n$-gram dictionary corresponding to the minimum of
	the desired $n$-gram length or the number of words we have already created.

	Perplexity works similarly, scanning the input list and computing the
	running sum or product probability, depending on whether we are using
	arbitrary precision or logarithmic arithmetic.  Because computing the
	perplexity for each word is simply a matter of looking up a tuple in the
	appropriate dictionary (possibly containing an unknown), this step is fairly
	rapid.  If we do not have the appropriate $n$-gram in the dictionary, we
	first try replacing the last word with an unknown, and, if that fails, we
	shorten the list of words until a match is found, which, at shortest, is a
	single unknown by itself.

	%\\

	The dictionary approach was simple and intuitive, but fairly slow; the
	generation of sentences spends the majority of its time on dictionary
	operations, and uses only moderate amounts of memory.  Much of the time in
	sentence generation is spent computing the list of $n$-gram options to
	choose from, since this requires, in our implementation, scanning through
	the entire $n$-gram dictionary to find matches for the $(n-1)$-gram.  If we
	precomputed this as we populated the dictionary, substantial time could be
	saved, although choosing a random element from that list still takes $O(n)$
	time.

\section*{Sentence Generation}

	For sentence generation, we did not include our unknown processing, and do
	not convert the first occurrence of each word to an unknown token.

	Each subsection contains two examples from the Shakespeare corpus and two
	from War and Peace.  These corpora were chosen because they were determined
	at rough glance to be the easiest to parse.

	\subsection*{Unsmoothed}

		\subsubsection*{Unigrams}

			\begin{quote}

			Bring in Take made the pass a Caesar in gold well leave

			Exeunt . hope , and Milan your A Will thy have Come the . .
			advantage , together your either to '

			He fell other along " therefore , and had about He chief shorten now
			him life . something

			I had transference shrugged man with you The lit . to lift the her
			had he . and have the hand individuals dragging the , place been to
			Natasha in inside . revelry hide nation room beaming eager all of
			the the , remounts to

			\end{quote}

		\subsubsection*{Bigrams}

			\begin{quote}

			Enter LYSANDER , and you lie dull actor on my speech , an ordinary
			pitch , Loyal and full of our tongues , see , That which physic to
			my horse for her youth and sealed bags Of base earth ; Though twenty
			several mistress will you have done To have found .

			Silence , who .

			The horses ' familiar to the army , envy and looked questioningly :
			"You won't be thus depriving him of the estate with his will send me
			forty thousand men in choosing her governess kept saying good-by .

			I have begun to Tushin and yet further battles , the man in the
			defense of that temple , they were standing beside whom you seen in
			the battle of dry branches of meeting it has brought the smile Berg
			with a mocking , and even impossible to be long paced angrily .

			\end{quote}

		\subsubsection*{Trigrams}

			\begin{quote}

			How - traitor ?

			The Greeks upon advice , Hath alter'd that good wisdom Whereof I
			know by their christen names , Unless thou tell her so .

			No , but it can't be helped It happens to the first to arrive ,
			Princess , I think .

			Rostopchin , coming to Petersburg , look out at night his feet .

			\end{quote}

	\subsection*{Smoothed}

		\subsubsection*{Unigram}

			\begin{quote}

			th ! come Lear things of the I

			me . I

			they its intimate just such But There the them said shout
			irresistible always heard none with and of dinner had To

			day good-natured Secondly same , joyful on talking the appeared
			advantage . unreasoning men by

			\end{quote}

		\subsubsection*{Bigram}

			\begin{quote}

			Hector .

			I arrest thee here , which time , that his fault , and blows .

			But my wife should have nobody and ask the enemy was too .

			Say it's farther .

			\end{quote}

		\subsubsection*{Trigram}

			\begin{quote}

			She would have made fair work !

			Come , my lord , I'll bring you to please his Majesty , Herod of
			Jewry dare not say he is much abus'd with tears thou keep'st command
			.

			All that he had no news from the chase .

			If I decline the honor of the bushes .

			\end{quote}

		Trigrams lead to more comprehensible sentences, which is to be expected,
		and bigrams are barely useful.  The unigrams also seemed to produce
		longer sentences; it's possible that trigrams and bigrams lead to a
		progression from words that are likely at the beginning of sentences to
		words that are likely at the end, and then to end the sentence, despite
		sentence ends being fairly infrequent.  Additionally, \textit{War and
		Peace} seems to produce longer sentences (particularly when observed
		over more iterations than can be reproduced comfortably here), probably
		because the sentences in the original are longer, so sentence ends are
		rarer.

		Our $n$-gram generator only treated sentence endings by adding a single
		symbol, `\#', and did not break the input into individual sentences, nor
		do we break the perplexity input into sentences in the same manner.

	\section*{Arbitrary Precision}

		One of the two extensions we chose was to implement the probability and
		perplexity calculations using arbitrary precision arithmetic (we also
		implemented it using logarithmic arithmetic).  This can lead to slower
		processing, but	guarantees that there is no data loss until we move into
		processing floats during exponentiation.

		We used python's +gmpy+ package to represent the probabilities as
		arbitrary rational numbers (python's integers are already
		arbitrary-precision), and +mpmath+ to perform the final exponentiation
		to calculate the perplexity (the $n$-th root).  The libraries could only
		compute non-integer powers of integers or floats, and integer powers
		were rounded to the nearest integer, and therefore, inaccurate, so
		(very high precision) floats provided the best results.  This final
		stage of computation of perplexity was the only point where imprecise
		mathematics was used.

		Our hypothesis was that the arbitrary precision math would reveal a
		small amount of error in computing the perplexity logarithmically.  Our
		results seem to indicate that any such error is very small (differing
		only in the last digit in measuring the perplexity of whole corpora),
		but that arbitrary-precision mathematics is much slower than floating
		point mathematics with logs when finding the product of many numbers.
		See \ref{runtimes}:

		\begin{table}
		\caption{Running times with arbitrary and floating-point precision in
		seconds.  Average of 10 trials, trained on the Shakespeare training
		corpus, perplexity calculated for Shakespeare training corpus}
		\label{runtimes}
		\begin{tabular}{l | r r}
			Operation & Arbitrary Precision & Logarithmic \\
			\hline
			Trigram tallying & 136.936949968 & 137.023640156\\
			Probability computation & 63.5648961067 & 60.1038038731 \\
			Perplexity computation & \textbf{990.322752953} & 65.2168021202 \\
		\end{tabular}
		\end{table}

	\section*{Arbitrary $n$-grams}

		Because the computations were general enough, we were able to easily
		abstract the code to produce count, smoothed count, and probability
		tables for arbitrary $n$-grams.  We also compute all of the smaller
		$n$-grams, which are required for generating or analyzing the beginning
		of a sequence of words, and for other computations.

		Large $n$-grams (particularly $n > 4$) tend to result in whole
		Shakespeare sentence fragments.  For example, ``O , my offence is rank ,
		it smells to heaven ; It hath the primal eldest curse upon't , A
		brother's murther !'' was generated by a hexagram, and is a direct quote
		from Hamlet, act \osn{3} scene \osn{3}.  This is a well-documented
		phenomenon, and is due to the fact that it's quite possible that the
		preceding $(n-1)$-gram appears only once in the entire corpus, so the
		sentence generator has no choice but to choose the next word in the
		quotation.

\section*{Experimental design}
We trained our model on the training data and then
ran two experiments to study the effect of \ngram length and
probability computation type on perplexity on the two corpora.
The two probability computation types were arbitrary precision
arithmetic and \loga.
For each experiment, we tested \ngram length of $n$=$1$ to $5$.

We hypothesized that longer $n$-grams would yield lower perplexity
and that the two probability computation types would yield similar perplexity.

\subsection*{Perplexity of test set}
We evaluated our model on the test set by computing
the perplexity of the test set. In addition to the variables
above, we ran this test with smoothing on and with smoothing off.
We hypothesized that smoothing would reduce perplexity.

\subsection*{Perplexity of generated sentences}
We also evaluated our model on sentences generated by our model.
We did this because sentence generation involves substantial
probability computations, where the difference in probability
computation type would be relevant.

For each combination of the intependent variables, we
generated 23 sentences from the \emph{War and Peace} corpus
and tested their perplexity. %ADD SHAKESPEARE

\section*{Results}

\subsection*{Pelplexity of test set}

\subsection*{Perplexity of generated sentences}
Perplexity of generated sentences decreased with increased
\ngram length

\begin{figure}
\begin{center}
<<echo=F,fig=T,height=9>>=
par(mfrow=c(2,1))
pp_gen.plot(results.shakespeare,'Shakespeare')
pp_gen.plot(results.tolstoy,'War and Peace')
@
\end{center}
\caption{\label{fig:gen}Perplexities of generated sentences by $n$-gram length and probability computation method for the two corpora.}
\end{figure}

\end{document}